### 2017-07-19 manuscript.Rmd
Almost finished first version of introduction

---
title: "Reversed motion direction illusion: influence of (spatially congruent) subsequent own-body motion on self-motion perception"
shorttitle: "Reversed motion direction illusion"
author:
  - name: First Author
    affiliation: 1
    corresponding: yes    # Define only one corresponding author
    address: Postal address
    email: my@email.com
  - name: Ernst-August Doelle
    affiliation: "1,2"
affiliation:
  - id: 1
    institution: Wilhelm-Wundt-Univerdddsity
  - id: 2
    institution: Konstanz Business School

abstract: |
  Enter abstract here (note the indentation, if you start a new paragraph).

author_note: |
  Complete departmental affiliations for each author (note the indentation, if you start a new paragraph).

  Enter author note here.

keywords: "keywords"

wordcount: X

class: man
lang: english
figsintext: yes
figurelist: no
tablelist: no
lineno: yes
bibliography:
  - references.bib
output: papaja::apa6_pdf
---

```{r message = FALSE, warning = FALSE}
library("papaja")
library("rstan")
library("brms")
library("tidyverse")
library("haven")
library("dplyr")

##### LOAD DATA #####

exp1 <- read_sav("SPSS Files Master/Exp 1/SPSS Data Files/final gelabelte Reaktionszeiten 11vp bereinigt, log transormiert.sav")

exp2 <- read_sav("SPSS Files Master/Exp 2/SPSS Data Files/gelabelte Reaktionszeiten 11 vp Exp 2 bereinigt, Log transformiert.sav")

exp3 <- read_sav("SPSS Files Master/Exp 3/SPSS Data Files/gelabelte Reaktionszeiten und Fehlerraten 12 vp Exp 3 bereinigt, Log transformiert.sav")

#### TRANSFORM DATA FOR ANALYSIS ####

## EXPERIMENT 1 ##

exp1 <- exp1 %>%
  dplyr::rename(subj = Vp,
                rt = ReactTimes,
                response = SubjectAnswer,
                correct = Richtigkeit,
                cue = Cue,
                soa = SOA)

exp1 <- exp1 %>% dplyr::mutate(logrt = log(rt),
                               response = ifelse(response == "R", 1, 0),
                               correct = ifelse(correct == 1, 1, 0),
                               dir2 = as.factor(ifelse(response == 1 & correct == 1, "right",
                                                       ifelse(response == 0 & correct == 0, "right", "left"))),
                               cue = factor(cue, labels = c("valid", "invalid", "neutral")),

                               dir1 = as.factor(ifelse(cue == "neutral", "forewards",
                                                       ifelse(dir2 == "right" & cue == "valid", "right",
                                                              ifelse(dir2 == "left" & cue == "invalid", "right", "left")))),
                               relevel(cue, ref = "neutral"),
                               soa = as.factor(soa),
                               subj = as.factor(subj)) %>%
  dplyr::select(subj, rt, logrt, response, correct, dir1, dir2, cue, soa)

## EXPERIMENT 2 ##

exp2 <- exp2 %>%
  dplyr::rename(subj = VP,
                rt = ReactTimes,
                response = SubjectAnswer,
                correct = Richtigkeit,
                cue = Cue,
                soa = SOA)

exp2 <- exp2 %>% dplyr::mutate(logrt = log(rt),
                               response = ifelse(response == "R", 1, 0),
                               correct = ifelse(correct == 1, 1, 0),
                               dir2 = as.factor(ifelse(response == 1 & correct == 1, "right",
                                                       ifelse(response == 0 & correct == 0, "right", "left"))),
                               cue = factor(cue, labels = c("valid", "invalid")),

                               dir1 = as.factor(ifelse(dir2 == "right" & cue == "valid", "right",
                                                              ifelse(dir2 == "left" & cue == "invalid", "right", "left"))),
                               soa = as.factor(soa),
                               subj = as.factor(subj)) %>%
  dplyr::select(subj, rt, logrt, response, correct, dir1, dir2, cue, soa)

## EXPERIMENT 3 ##

exp3 <- exp3 %>%
  dplyr::rename(subj = VP,
                rt = ReactTimes,
                response = SubjectAnswer,
                correct = Richtigkeit,
                cue = Cue,
                soa = SOA)

exp3 <- exp3 %>% dplyr::mutate(logrt = log(rt),
                               response = ifelse(response == "R", 1, 0),
                               correct = ifelse(correct == 1, 1, 0),
                               dir2 = as.factor(ifelse(response == 1 & correct == 1, "right",
                                                       ifelse(response == 0 & correct == 0, "right", "left"))),
                               cue = factor(cue, labels = c("valid", "invalid")),

                               dir1 = as.factor(ifelse(dir2 == "right" & cue == "valid", "right",
                                                       ifelse(dir2 == "left" & cue == "invalid", "right", "left"))),
                               soa = as.factor(soa),
                               subj = as.factor(subj)) %>%
  dplyr::select(subj, rt, logrt, response, correct, dir1, dir2, cue, soa)

### SOME DESCRIPTIVES ###

# Add Experiment variable 
exp1$experiment <- 'exp1'
exp2$experiment <- 'exp2'
exp3$experiment <- 'exp3'

#Redfinde Subject Variable
levels(exp1$subj) <- 1:length(unique(exp1$subj))
levels(exp2$subj) <- length(unique(exp1$subj))+1:length(unique(exp2$subj))
levels(exp3$subj) <- length(unique(exp2$subj))+length(unique(exp2$subj))+1:length(unique(exp3$subj))

# Put all three experiments into one df.
exp_all <- rbind(exp1, exp2, exp3)
exp_all$experiment <- factor(exp_all$experiment)

# Get mean correctness for all experiments

exp_all_correct <- exp_all %>% 
  group_by(subj, cue, soa, experiment) %>%
  summarise(meanCorrect = mean(correct))

# Rename Variables for nice figure

levels(exp_all_correct$cue) <- c('congruent', 'incongruent', 'neutral')
levels(exp_all_correct$soa) <- c('50', '100', '200', '600')
levels(exp_all_correct$experiment) <- c('Experiment 1', 'Experiment 2', 'Experiment 3')

# Get mean correctness for all experiments

  exp_all_rt <- exp_all %>% 
    group_by(subj, cue, soa, experiment, correct) %>%
    summarise(meanLogRT = mean(log(rt)))

# Rename Variables for nice figure

  levels(exp_all_rt$cue) <- c('congruent', 'incongruent', 'neutral')
  levels(exp_all_rt$soa) <- c('50', '100', '200', '600')
  levels(exp_all_rt$experiment) <- c('Experiment 1', 'Experiment 2', 'Experiment 3')
  exp_all_rt$correct <- factor(exp_all_rt$correct)
  levels(exp_all_rt$correct) <- c('Wrong', 'Correct')
```

# Abstract 
Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.

Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto.

Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet.

# Introduction
The malleable nature of human perception is best observed in sensory illusions. One of such illusions is known as motion aftereffect (MAE): Classically, in the MAE a prolonged exposure to moving visual stimulus leads to the temoporarily illusory perception of motion in the other opposite direction of a subsequent static stimulus. Famously described by Adams as the waterfall illusion [@adams_account_1834], the MAE has been extensively studied in the visual modality. Albeit their simplicity, MAEs are a powerful tool to investigate underlying neural and computational mechanisms of, mainly visual, motion perception [@anstis_motion_1998; @cuturi_optic_2014; @konkle_motion_2009]. Dominating explanation models rely on selective adaptation of motion-sensitive neurons, mostly demonstrated in the visual system. In such models, a prolonged exposition to a specific motion leads to a reduced firing rate and responiveness of those neurons. Assuming that distinct populations of neurons code different directions which are constantly compared, a prolonged stimulation with one direction and thus resulting decreased firing rate would result in the perception apparent motion in the opposite direction [@anstis_motion_1998].

However, MAEs are not limited to the visual modality and similar effects have also been observed in the auditory [@bestelmeyer_auditory_2010; @latinus_perceptual_2012; @reinhardt-rutland_spectrally_2001], proprioceptive [@seizova-cajic_proprioceptive_2007] and recently also vestibular modality [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012]. The modality independent nature of the MAE is further highlighted in crossmodal aftereffects. While there are multisensory effects in the visuo-auditory [@berger_auditory_2016; @ehrenstein_cross-modal_1996] and visuo-tactile domain [@konkle_motion_2009], for the present study visuo-vestibular aftereffects are more relevant.

Studies on visuo-vestibular aftereffects make use of a special kind of visual stimulation, i.e. vection. In vection, visual motion stimuli that simulate the retinal stimulation generated body movements elicit the illusory feeling of self motion [@brandt_motion_1974]. In this context it has been shown, that vection also results in vection after effects [@patterson_modeling_2009; @seno_vection_2010]. Additionally, vestibular information seems to modulate the aftereffects induced by vection [@wallach_compensation_1975]. In a seminal study, Cuturi and MacNeilage Thus, it has been investigated whether the visuo-vestibular coupling is able to produce visuo-vestibular MAE [@cuturi_optic_2014] . Taken together, these studies point out that there are aftereffects for self-motion as well, which are found for the visual system alone (vection), the vestibular system alone [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012] and their combination [@cuturi_optic_2014]. Thus, MAEs seem to be a sensory modality independent phenomenon and an explanatory models limited to lower-level visual areas fail to grasp the crossmodal nature of the sensory motion aftereffect.

For the vestibular system, possible MAEs are highly interesting and important as the vestibular organs track the position of the self relative to space and are thus highly involved in the perception of self motion. Anecdotal evidence for such vestibular MAEs are found in the 'Gillingham illusion' [@ercoline_post-roll_2000], in which pilots tend to counter previous roll motions with a roll motion in the same direction. Driven by this phenomenon, Crane and colleagues found such aftereffects for fore-aft translation [@crane_foreaft_2012], roll [@crane_roll_2012] and yaw rotation [@coniglio_human_2014]. This also highlights that vestibular aftereffects can be evoked in the semicircular canals as well as in the otoliths, giving further evidence for the crossmodal nature of motion aftereffects.

Taken together, these findings contribute to the idea that there is a centre for self-motion perception that integrates information from different modalities [@seno_vection_2010]. However, in the vestibular system an important distinction has to be made between the perception of rotation, which is coded in the semicircular canals, and the perception of translation, which is coded in the otoliths, although their signals are combined at an early processing stage [@crane_perception_2016]. Thus, humans percept of selfmotion is multisensory even within the vestibular system. In every-day life those two signals are rarely perceived alone, which makes crossmodal aftereffects within the vestibular system more interesting.

In the present study, we set out to explore this idea even more with data from three experiments that tried to shift attention with natural vestibular stimulation consisting of a prior yaw rotation or inertial translation to facilitate the discrimination of the direction of a subsequent congruent or incongruent yaw rotation or inertial translation. Thus, in experiment one a inertial translation was always followed by a yaw rotation in the same or opposite direction, and in experiment two and three a yaw rotation was followed by a inertial translation in the same or opposite direction. Participants' task was always to indicate the perceived direction of the second motion as fast as possible. Thus, the three experiments investigate whether there is a crossmodal aftereffect within the vestibular system and whether it is unilateral.

The structure of the three experiments with the discrimination tasks is ideal to study the perceptual decision making process. As far as we know, the decision making process has never been investigated more closely in vestibular motion aftereffects. However, a joint analysis of response and response time would provide more detailed information about the dynamics of the underlying process. A commonly used model for a joint analysis is the Drift Diffusion Model (DDM) [@ratcliff_diffusion_2008]. In the DDM it is assumed that the decision making process relies on the acculumation of noisy sensory evidence over time. A response is given when a threshold for one of the two possibilities is reached. Several parameters describe the decision making process: The drift rate stands for the speed of evidence accumulation for one of the two response options, the boundary separation parameter serves as the threshold for the needed evidence to make response, the non deicision time parameter covers the portion within a response not involved in the deicision making process (e.g. motor execution) and the starting point indicates a point within the boundary from which evidence accumulation starts. Other studies within the field of perceptual decision making [@ellis_vestibular_2017; @wiech_influence_2014] used DDM to disentagle influence of prior information on parameters of the decision making process. From another point of view, in the MAE a first movement introduces prior information which influences the evaluation of the second motion. A DDM allows us to investigate altered sensory processing and a perceptual decision-making bias separately. So far MAE or TAE did not consider those two different biases.

# Experiment 1

## Methods

### Participants

Twelve healthy participants participated in the first experiment (four female, mean age 27, range 24-30 years). All participants were right handed according to a Gemran version of the handedness questionnaire by Chapman and Chapman [@chapman_measurement_1987]. One of the participants received a course credit for participation. None of the participants reported a history of relevant neurological, vestibular or attentional disorders. The study was approved by the ethics committeeof the University of Bern and all participants gave written informed consent prior to the experiment in accordance with the Declaration of Helsinki.

### Motion stimuli

A six degree of freedom motion base (6DOF2000E, Moog Inc., East Aurora, NY) and in-house software were used to generate the motion stimuli. Cue (antecedent) stimuli consisted of translations with single cycle sinusoidal acceleration and a frequency $f$ of 5 Hz ($a(t) = A\sin(2 \pi ft) = A\sin(\frac{2 \pi t}{T})$, $T = \frac{1}{f}$) along the y-axis (left/right) and the x-axis (forth/back) as used in previous studies (see e.g. Grabherr et al. 2008; Benson et al. 1989; Crane 2012). Acceleration amplitude (A) was set to $0.25 \frac{m}{s^2}$, i.e. a peak velocity of $0.016 \frac{m}{s}$ ($v_{max} = \frac{AT}{\pi}$) and a displacement of 0.0016 m ($\Delta p = \frac{AT^2}{2\pi}$). For the target (subsequent) stimuli the same acceleration profile as for the cue stimuli was used but with yaw rotations about an earth-vertical axis (left/right). The acceleration amplitude was set to $24 \frac{deg}{s^2}$ ($v_{max} = 1.53 \frac{deg}{s}, \Delta p = .153 deg$).

### Experimental design

Within each participant Cue Congruency (three levels: congruent, incongruent and neutral) and interstimulus interval (ISI, four levels: 50 ms, 100 ms, 200 ms and 600 ms) were manipulated. In congruent trials y-axis translations were followed by yaw rotations to the corresponding side. In incongruent trials y-axis translations preceeded yaw rotations to the opposite side. Neutral trials consisted of a translation along the x-axis and a subsequent yaw rotation. The combination of the two factors resulted in 32 motion sets. Every motion set was repeated six times per participant. In-house software based on labVIEW (CITE) was used to record participants' response and reaction times.

### Experimental procedure

Participants were seated upright in a car seat with a five-point harness which was mounted on the motion base. Their head was centered and fixated with a helmet. The experiments were conducted in the dark and participants were blindfolded to prevent the processing of visual cues. White noise was delivered to in-ear headphones at approximately 60 dB to mask sounds from the motion base. Participants were instructed to indicate the perceived direction of the second motion as fast as possible by pressing the corresponding button in either their left or right hand. In case of uncertainty, participants had to make their best guess. The first motion was described as a time varying warning cue, which would indicate the beginning of the second motion but not predict its direction. Practice trials were administered until participants understood the task. Once a participant was comfortable with the task, motion sets were presented in three blocks. The length of the breaks between blocks were determined by participants. In total, 192 trials were randomly presented (48 congruent, 48 incongruent, 96 neutral). A trial consisted of a cue stimulus (200 ms), an ISI (50-600 ms), a target stimulus (200 ms), time to respond (max. 2500 ms), return to the origin (1100 ms) and an intertrial interval of about 1000 ms.

### Data analysis

All analyses and figures were computed in R [@R-base] using the ggplot2 [@R-ggplot2], brms [@R-brms] and rstan [@R-rstan] packages. 
To describe data from experiment 1 in more detail, a Bayesian multilevel logistic regression with treatment coded factors Cue Congruence (congruent, incongruent, neutral) and ISI (50ms, 100ms, 200ms, 600ms) was calculated. For the reaction times a Bayesian multilevel regression with a ex-Gaussian likelihood function and factors accuracy (correct, wrong), Cue Congruence (congruent, incongruent, neutral) and ISI (50ms, 100ms, 200ms, 600ms) was computed. Priors were set to default in brms.

Several Drift Diffusion Models implented in brms were used to fit the parameters of interest. A drift diffusion model "simulates" a decision making process with the following parameters: drift rate, boundary separation, starting point and non-decision time.  The upper boundary was coded as decision for 'right motion' while the lower boundary was a 'left motion'. Accuracy was not inlcuded in the models but can be infered from the beta parameters.

## Results

## Discussion

# Experiment 2

The results of Experiment 1 led to the question whether the consistent directional misperception of the second motion still exists when a rotation (cue) is preceding a translation (target). The sequence of the two motions was therefore interchanged so that a left/right yaw rotation preceded a left/right translation (while the motion trajectories stayed otherwise the same). Only congruent and incongruent cues were presented, since no rotations could serve as “neutral” cues for lateral translations in the horizontal plane. A pitch or roll rotation in an upright participant would lead inevitably to a tilted/skewed subsequent motion.

## Methods

### Participants

For experiment 2, twelve new participants were recruited. One participant pressed the left button in 92% of all trials and was excluded from the study. The remaining 11 participants (nine female, mean age 25, range 21-38 years). All participants were right handed according to a German version of the handedness questionnaire by Chapman and Chapman [@chapman_measurement_1987]. None of the participants reported a history of relevant neurological, vestibular or attentional disorders. The study was approved by the ethics committeeof the University of Bern and all participants gave written informed consent prior to the experiment in accordance with the Declaration of Helsinki.

### Motion stimuli

Cue stimuli consisted of yaw rotations about an earth-vertical axis (left/right). They consisted of single-cycles of sinusoidal acceleration at a frequency of 5 Hz. Acceleration amplitude was $24 \frac{deg}{s^2}$ ($v_{max} = 1.53 \frac{deg}{s}, \Delta p = .0153 deg$). Target stimuli were translations along the y-axis (left/right). They consisted of the same acceleration profile as the cue stimuli. Frequency was again 5 Hz. Acceleration amplitude was $0.25 \frac{m}{s^2}$ ($v_{max} = .016 \frac{m}{s}, \Delta p = .0016 m$).

### Experimental design

Within each participant Cue Congruency (three levels: congruent, incongruent) and interstimulus interval (ISI, four levels: 50 ms, 100 ms, 200 ms and 600 ms) were manipulated. In congruent trials yaw rotation were followed by translations along the y-axis to the corresponding side. In incongruent trials yaw rotations were followed by y-axis translations to the opposite side. The combination of the two factors resulted in 16 motion sets. Every motion set was repeated 12 times per participant. Again, participants' response and reaction times were recorded.

### Experimental procedure

Experimental procedure was similar to the one in Experiment 1. Only the type of motion sets presented was different. In this experiment, 192 motion sets were randomly presented (96 congruent, 96 incongruent).

### Data analysis

All analyses were performed in R [@R-base] using the brms [@R-brms] and rstan [@R-rstan] packages.

## Results
```{r figDescriptiveCorrect, fig.cap='The means for all conditions and their accuracy'}
exp_all_correct %>%
  ggplot(aes(x = soa, y = meanCorrect, shape = cue, color = cue)) + 
  geom_hline(yintercept = .5, linetype = 'dashed') +
  stat_summary(fun.y = mean, geom = 'point') + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar") + 
  #geom_boxplot() +
  coord_cartesian(ylim = c(0,1), expand = 0) + 
  scale_y_continuous(limits = c(0,1)) + 
  labs(x = 'Inter-Stimulus-Intervall', y = 'Accuracy', 
       title = 'Accuracy over all Conditions', 
       color = 'Motion Congruence', 
       shape = 'Motion Congruence') + 
  ylab('Accuracy') + 
  xlab('Inter-Stimulus-Intervall') +
  facet_grid(~experiment) +
  theme_bw(base_size = 12) +
  theme(plot.title = element_text(size = 12, face="bold", hjust = 0.5), 
        axis.title = element_text(size = 12, face = 'bold'), 
        legend.title = element_text(size = 12, face = 'bold'), 
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())
```

```{r figDescriptiveRT}         
na.omit(exp_all_rt) %>%
    ggplot(aes(x = soa, y = meanLogRT, shape = cue, color = cue, linetype = cue)) + 
    #stat_summary(fun.y = mean, geom = 'point') + 
    #stat_summary(fun.data = mean_cl_normal, geom = "errorbar") + 
    geom_boxplot() +
    labs(x = 'Inter-Stimulus-Intervall', y = 'Accuracy', 
         title = 'Reaction Times', 
         color = 'Motion Congruence', 
         shape = 'Motion Congruence',
         linetype = 'Motion Congruence') + 
    ylab('Mean Log Reaction Time') + 
    xlab('Inter-Stimulus-Intervall') +
    facet_grid(experiment ~ correct) +
    theme_bw(base_size = 12) +
    theme(plot.title = element_text(size = 12, face="bold", hjust = 0.5), 
          axis.title = element_text(size = 12, face = 'bold'), 
          legend.title = element_text(size = 12, face = 'bold'), 
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
```

## Discussion

# Experiment 3

In Experiment 2 motion stimuli were the same as in Experiment 1, simply in a different order. The cue stimuli in Experiment 1 were stronger in terms of intensity compared to the target stimuli. Furthermore, because their order was switched in Experiment 2 without adaption of their intensities, the absence of the strong effect for error rates found in Experiment 1 can easily be explained. However, within the data of Experiment 2, there were two participants whose error rates data suggested that they consistently misinterpreted the direction of congruent cued target stimuli. It was thus hypothesized that the effect would reappear if the intensities of the stimuli in terms of distance to thresholds were in the same ratio as in Experiment 1. For this third experiment the rotation about an earth-vertical axis was chosen at an intensity 3.6 times above the threshold found by Grabherr et al. [@grabherr_vestibular_2008] and the horizontal translation was chosen at an intensity 2.6 times above the threshold found by Lim et al. (in preparation).

## Methods

### Participants
Twelve new participants were recruited for the third experiment. Two participants had to abort the experiment. The remaining ten participants (four female, mean age 29, range 23-56 years). All participants were right handed according to a German version of the handedness questionnaire by Chapman and Chapman [@chapman_measurement_1987]. None of the participants reported a history of relevant neurological, vestibular or attentional disorders. The study was approved by the ethics committeeof the University of Bern and all participants gave written informed consent prior to the experiment in accordance with the Declaration of Helsinki.

### Motion stimuli

Cue stimuli were rotations about an earth-vertical axis (right/left) with single cycle sinusoidal acceleration and a frequency of 5 Hz ($a(t) = A \sin (2 \pi ft) = A \sin (\frac{2\pi t}T)$). Acceleration amplitude was set to $33 \frac{deg}{s^2}$ ($v_{max} = 2.10 \frac{deg}{s}, \Delta p = 0.021 deg$). Target stimuli consisted of translations in the y-axis (right/left) and the same acceleration profile as the cue stimuli. Acceleration amplitude was set to $0.18 \frac{m}{s^2}$ ($v_{max} = .011 \frac{m}{s}, \Delta p = .0011 m$).

### Experimental design and procedure

Design and experimental procedure were identical to Experiment 2.

### Data analysis

All analyses were performed in R [@R-base] using the brms [@R-brms] and rstan [@R-rstan] packages

## Results

## Discussion

# General Discussion

- ISIs are longer than compared to Crane. At their ISI our effect almost disapears. They argue that more reliable at 1 or 3 sec than 0.5 sec, longer latency? In contrast with our result, but they used different paradigm, weaker stimuli or Maybe not same underlying mechanism? But in Yaw Rotation (Coniglio and Crane, 2014) showed similar effect: with increasing ISI aftereffect became weaker, although they used longer ISI than we did.

- In the visual modality it is usually shown after several seconds of stimulation but has also been shown for very short stimulation of 320ms (kanai verstraten 2005)
- One possibility: ISI to short and sampling rate of vestibular system is not able to discriminate. BUT: movements are strong and stimulated vestibular organs changes from first to second motion. 
- also this does not explain why for congruent motion participants chose opposite direction if they would stick to first motion. 
- Extra proprioceptive input cannot be exluded especially given the strong nature of the motions (Seizova-Cajic, 2007)

However, self-motion perception is more complex, and other systems such as the proprioception are also involved. Thus, the visuo-tactile motion aftereffect could provide further evidence for such a higher level centre for self motion.
## Future Research

Given the original nature of the experimental idea, the designs differs quite a lot to other experiments in this domain. They mostly made use of methods from psychophysics with adapte test stimuli, while in the present experiment constant motion stimuli were used.

Of special interest: Visuo-vestibular integration. This is integration is highly important to disambiguate self and object motion: Can we find such an aftereffect also for visuo-vestibular integration? Oculus?

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}