### 2017-07-27 manuscript.Rmd
first long version of manuscript

---
title: "Motion Aftereffects in the vestibular system: influence of (spatially congruent) subsequent own-body motion on self-motion perception"
shorttitle: "Reversed motion direction illusion"
author:
  - name: "Roman B. Di Francesco"
    affiliation: "1"
  - name: "Andrew W. Ellis"
    affiliation: "1"
  - name: "Luzia Grabherr"  
    affiliation: "1"
  - name: "Gianluca Macauda"
    affiliation: "1"
    corresponding: yes    # Define only one corresponding author
    address: "Departement of Psychology, University of Bern, Fabrikstrasse 8, 3012 Bern, Switzerland"
    email: "gianluca.macauda@psy.unibe.ch"
  - name: "Fred W. Mast"
    affiliation: "1"
affiliation:
  - id: "1"
    institution: "Departement of Psychology, University of Bern"

author_note: >
  Complete departmental affiliations for each author (note the indentation, if you start a new paragraph).

abstract: >
  Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem ipsum dolor sit amet. Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse molestie consequat, vel illum dolore eu feugiat nulla facilisis at vero eros et accumsan et iusto.

keywords: "keywords"

wordcount: X

class: man
lang: english
figsintext: yes
figurelist: no
tablelist: no
lineno: yes
bibliography:
  - references.bib
output: papaja::apa6_pdf
---

```{r message = FALSE, warning = FALSE}
library("papaja")
library("rstan")
library("brms")
library("tidyverse")
library("haven")
library("dplyr")

##### LOAD DATA #####

exp1 <- read_sav("SPSS Files Master/Exp 1/SPSS Data Files/final gelabelte Reaktionszeiten 11vp bereinigt, log transormiert.sav")

exp2 <- read_sav("SPSS Files Master/Exp 2/SPSS Data Files/gelabelte Reaktionszeiten 11 vp Exp 2 bereinigt, Log transformiert.sav")

exp3 <- read_sav("SPSS Files Master/Exp 3/SPSS Data Files/gelabelte Reaktionszeiten und Fehlerraten 12 vp Exp 3 bereinigt, Log transformiert.sav")

#### TRANSFORM DATA FOR ANALYSIS ####

## EXPERIMENT 1 ##

exp1 <- exp1 %>%
  dplyr::rename(subj = Vp,
                rt = ReactTimes,
                response = SubjectAnswer,
                correct = Richtigkeit,
                cue = Cue,
                soa = SOA)

exp1 <- exp1 %>% dplyr::mutate(logrt = log(rt),
                               response = ifelse(response == "R", 1, 0),
                               correct = ifelse(correct == 1, 1, 0),
                               dir2 = as.factor(ifelse(response == 1 & correct == 1, "right",
                                                       ifelse(response == 0 & correct == 0, "right", "left"))),
                               cue = factor(cue, labels = c("valid", "invalid", "neutral")),

                               dir1 = as.factor(ifelse(cue == "neutral", "forewards",
                                                       ifelse(dir2 == "right" & cue == "valid", "right",
                                                              ifelse(dir2 == "left" & cue == "invalid", "right", "left")))),
                               relevel(cue, ref = "neutral"),
                               soa = as.factor(soa),
                               subj = as.factor(subj)) %>%
  dplyr::select(subj, rt, logrt, response, correct, dir1, dir2, cue, soa)

## EXPERIMENT 2 ##

exp2 <- exp2 %>%
  dplyr::rename(subj = VP,
                rt = ReactTimes,
                response = SubjectAnswer,
                correct = Richtigkeit,
                cue = Cue,
                soa = SOA)

exp2 <- exp2 %>% dplyr::mutate(logrt = log(rt),
                               response = ifelse(response == "R", 1, 0),
                               correct = ifelse(correct == 1, 1, 0),
                               dir2 = as.factor(ifelse(response == 1 & correct == 1, "right",
                                                       ifelse(response == 0 & correct == 0, "right", "left"))),
                               cue = factor(cue, labels = c("valid", "invalid")),

                               dir1 = as.factor(ifelse(dir2 == "right" & cue == "valid", "right",
                                                              ifelse(dir2 == "left" & cue == "invalid", "right", "left"))),
                               soa = as.factor(soa),
                               subj = as.factor(subj)) %>%
  dplyr::select(subj, rt, logrt, response, correct, dir1, dir2, cue, soa)

## EXPERIMENT 3 ##

exp3 <- exp3 %>%
  dplyr::rename(subj = VP,
                rt = ReactTimes,
                response = SubjectAnswer,
                correct = Richtigkeit,
                cue = Cue,
                soa = SOA)

exp3 <- exp3 %>% dplyr::mutate(logrt = log(rt),
                               response = ifelse(response == "R", 1, 0),
                               correct = ifelse(correct == 1, 1, 0),
                               dir2 = as.factor(ifelse(response == 1 & correct == 1, "right",
                                                       ifelse(response == 0 & correct == 0, "right", "left"))),
                               cue = factor(cue, labels = c("valid", "invalid")),

                               dir1 = as.factor(ifelse(dir2 == "right" & cue == "valid", "right",
                                                       ifelse(dir2 == "left" & cue == "invalid", "right", "left"))),
                               soa = as.factor(soa),
                               subj = as.factor(subj)) %>%
  dplyr::select(subj, rt, logrt, response, correct, dir1, dir2, cue, soa)

### SOME DESCRIPTIVES ###

# Add Experiment variable 
exp1$experiment <- 'exp1'
exp2$experiment <- 'exp2'
exp3$experiment <- 'exp3'

#Redfinde Subject Variable
levels(exp1$subj) <- 1:length(unique(exp1$subj))
levels(exp2$subj) <- length(unique(exp1$subj))+1:length(unique(exp2$subj))
levels(exp3$subj) <- length(unique(exp2$subj))+length(unique(exp2$subj))+1:length(unique(exp3$subj))

# Put all three experiments into one df.
exp_all <- rbind(exp1, exp2, exp3)
exp_all$experiment <- factor(exp_all$experiment)

# Get mean correctness for all experiments

exp_all_correct <- exp_all %>% 
  group_by(subj, cue, soa, experiment) %>%
  summarise(meanCorrect = mean(correct))

# Rename Variables for nice figure

levels(exp_all_correct$cue) <- c('congruent', 'incongruent', 'neutral')
levels(exp_all_correct$soa) <- c('50', '100', '200', '600')
levels(exp_all_correct$experiment) <- c('Experiment 1', 'Experiment 2', 'Experiment 3')

# Get mean correctness for all experiments

  exp_all_rt <- exp_all %>% 
    group_by(subj, cue, soa, experiment, correct) %>%
    summarise(meanLogRT = mean(log(rt)))

# Rename Variables for nice figure

  levels(exp_all_rt$cue) <- c('congruent', 'incongruent', 'neutral')
  levels(exp_all_rt$soa) <- c('50', '100', '200', '600')
  levels(exp_all_rt$experiment) <- c('Experiment 1', 'Experiment 2', 'Experiment 3')
  exp_all_rt$correct <- factor(exp_all_rt$correct)
  levels(exp_all_rt$correct) <- c('Wrong', 'Correct')

plot_correct <- exp_all_correct %>%
  ggplot(aes(x = soa, y = meanCorrect, shape = cue, color = cue)) + 
  geom_hline(yintercept = .5, linetype = 'dashed') +
  stat_summary(fun.y = mean, geom = 'point') + 
  stat_summary(fun.data = mean_se, geom = "errorbar") + 
  #geom_boxplot() +
  coord_cartesian(ylim = c(0,1), expand = 0) + 
  scale_y_continuous(limits = c(0,1)) + 
  labs(x = 'Inter-Stimulus-Intervall', y = 'Accuracy', 
       title = 'Accuracy over all Conditions', 
       color = 'Motion Congruence', 
       shape = 'Motion Congruence') + 
  ylab('Accuracy') + 
  xlab('Inter-Stimulus-Intervall') +
  facet_grid(~experiment) +
  theme_bw(base_size = 12) +
  theme(plot.title = element_text(size = 12, face="bold", hjust = 0.5), 
        axis.title = element_text(size = 12, face = 'bold'), 
        legend.title = element_text(size = 12, face = 'bold'), 
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())

rt_times_plot <- na.omit(exp_all_rt) %>%
    ggplot(aes(x = soa, y = meanLogRT, shape = cue, color = cue, linetype = cue)) + 
    #stat_summary(fun.y = mean, geom = 'point') + 
    #stat_summary(fun.data = mean_cl_normal, geom = "errorbar") + 
    geom_boxplot() +
    labs(x = 'Inter-Stimulus-Intervall', y = 'Accuracy', 
         title = 'Reaction Times', 
         color = 'Motion Congruence', 
         shape = 'Motion Congruence',
         linetype = 'Motion Congruence') + 
    ylab('Mean Log Reaction Time') + 
    xlab('Inter-Stimulus-Intervall') +
    facet_grid(experiment ~ correct) +
    theme_bw(base_size = 12) +
    theme(plot.title = element_text(size = 12, face="bold", hjust = 0.5), 
          axis.title = element_text(size = 12, face = 'bold'), 
          legend.title = element_text(size = 12, face = 'bold'), 
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())

load('figures_ddm.Rda')
load('figures_ddm_separate.Rda')
```
# Introduction
The malleable nature of human perception is best observed in sensory illusions. One of such illusions is known as motion aftereffect (MAE): Classically, in the MAE a prolonged exposure to a moving visual stimulus leads to the temporary illusory perception of motion to the opposite direction of a subsequent static stimulus. Famously described by Adams as the waterfall illusion [@adams_account_1834], the MAE has evolved into an extremely interesting paradigm for vision science. Albeit their simplicity, MAEs are a powerful tool to investigate underlying neural and computational mechanisms of motion perception [@anstis_motion_1998; @cuturi_optic_2014; @konkle_motion_2009]. In visual sciences, the dominating explanation models rely on a selective adaptation mechanism of motion-sensitive neurons. In such models, a prolonged exposition to a specific motion subsequently leads to a reduced firing rate and responsiveness of those neurons. Assuming that distinct populations of neurons code different directions which are constantly compared, a prolonged stimulation with one direction would result in a decreased firing rate and thus in the perception of apparent motion to the opposite direction of static patterns [@anstis_motion_1998].

However, MAEs are not limited to the visual modality and similar effects have also been observed in the auditory [@bestelmeyer_auditory_2010; @latinus_perceptual_2012; @reinhardt-rutland_spectrally_2001], proprioceptive [@seizova-cajic_proprioceptive_2007] and recently also vestibular sense [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012]. The multisensory nature of the MAE is further highlighted in cross-modal aftereffects. Multisensory effects have been observed in the visuo-auditory [@berger_auditory_2016; @ehrenstein_cross-modal_1996] and visuo-tactile domain [@konkle_motion_2009], but for the present study we will focus visuo-vestibular MAEs.

Studies on visuo-vestibular MAEs make use of a special kind of visual stimulation, i.e. vection. In vection, visual motion stimuli elicit the illusory feeling of self-motion [@brandt_motion_1974]. In this context it has been shown, that vection results in MAEs regarding the self [@patterson_modeling_2009; @seno_vection_2010]. Additionally, vestibular information seems to modulate the aftereffects induced by vection [@wallach_compensation_1975]. In their seminal study, @cuturi_optic_2014 observed that sustained optic flow induced aftereffects in a subsequent vestibular motion detection task and thus demonstrated a cross-modal MAE between the visual and vestibular senses that may rely on different neurophysiological mechanisms than the visual MAE. Taken together, these studies point out that aftereffects are not limited to visual motion in general but also specific for self-motion, which are found in vection paradigms, in the vestibular system alone [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012] and in the combination of the visual and vestibular senses [@cuturi_optic_2014]. Thus, MAEs seem to be a sensory modality independent phenomenon and an explanatory models limited to lower-level visual areas fail to grasp the cross-modal nature of the sensory motion aftereffect.

For the vestibular system, possible MAEs are highly interesting and important as the vestibular organs track the position of the self relative to space and are thus highly involved in the perception of self motion. Anecdotal evidence for such vestibular MAEs is reported in the 'Gillingham illusion' [@ercoline_post-roll_2000], a phenomenon in which pilots tend to counter previous roll motions with a roll motion in the same direction. Driven by this illusion, @crane_foreaft_2012 and colleagues found such aftereffects for fore-aft translations, roll [@crane_roll_2012] and yaw rotations [@coniglio_human_2014]. This also highlights that vestibular aftereffects include both the semicircular canals as well as in the otoliths. Those findings provide further evidence for the cross-modal nature of motion aftereffects.

Taken together, these findings also contribute to the idea that there is a center for self-motion perception that integrates information from different modalities [@seno_vection_2010]. However, in the vestibular system a distinction has to be made between the perception of rotation, which is coded in the semicircular canals, and the perception of translation, which is coded in the otoliths, although their signals are combined at an early processing stage [@crane_perception_2016; @carriot_integration_2015; @dickman_vestibular_2002]. Thus, humans percept of self-motion is multisensory even within the vestibular system. In every-day life those two signals are rarely perceived separately, which makes the possibilty of cross-modal aftereffects within the vestibular system even more interesting.

In the present study, we set out to explore this idea with data from three experiments (unpublished data) that tried to shift attention with natural vestibular stimulation consisting of a prior yaw rotation or translation to facilitate the discrimination of the direction of a subsequent spatially congruent or incongruent yaw rotation or translation. Thus, in experiment 1 a translation was always followed by a yaw rotation in the same or opposite direction, and in experiment 2 and 3 a yaw rotation was followed by a translation to the same or opposite direction. The participants' task was always to indicate the perceived direction of the second motion as fast as possible. We used those three experiments to investigate whether there is a cross-modal MAE within the vestibular system and whether it is unilateral.

The design of the three experiments with the discrimination tasks is ideal to study the perceptual decision making process. As far as we know, the decision making process has never been investigated in MAEs. However, a joint analysis of response and response time would provide more detailed information about the dynamics of the underlying sensory processing.

A commonly used cognitive model for a joint analysis is the Drift Diffusion Model (DDM) [@ratcliff_diffusion_2008]. In the DDM it is assumed that the decision making process relies on the accumulation of noisy sensory evidence over time. A response is given when a threshold for one of the two possibilities is reached. Several parameters specify the decision making process: The drift rate describes the speed of evidence accumulation for one of the two response options. The starting point indicates a point within the boundary from which evidence accumulation starts. Parameters of interest for the current study are the drift rate and starting point. Other important parameters of the DDM include the boundary separation parameter, which serves as the threshold for the needed evidence to make decision and the non decision time parameter that covers the portion within a response not involved in the decision making process (e.g. motor execution).

Other studies within the field of perceptual decision making [@ellis_vestibular_2017; @wiech_influence_2014] used a DDM to investigate the influence of prior information on different parameters of the decision making process and has recently been proposed as model to study the dynamics of vestibular decision making [@merfeld_dynamics_2016] Applied to MAEs, a first motion stimulation (in our case vestibular stimulation) should influence the judgment of the subsequent motion discrimination task by introducing a bias. However, it remains to be seen whether such a bias shifts the starting point to the opposite direction, which would mean that a first motion alters the perceptual decision making process, or whether the first motion affects the drift rate and thus changes the accumulation process, which would be indicative for altered sensory processing [@wiech_influence_2014]. A combination of both biases is possible as well. Those three possbilities are discussed more extensively in a simplified depicition in Figure \@ref(fig:ddms)).

The goal of this study is twofold: First, we want to replicate the previously found MAE in the vestibular modality. However, here it tested whether a stimulation of the otoliths can evoke MAE for the semicircular canals and vice versa. This would further highlight the cross-modal nature of MAEs. Second and based on the MAE, it is investigated whether the introduced bias for the opposite direction is more likely due to changes in sensory processing or a more general cognitive strategy with the help of a joint analysis of response times and responses in a DDM.

```{r ddms, fig.cap = 'Three possibilites how the first motion could influence the discrimination of the second motion to elicit a MAE. On the left, the perceptual decision making process would be changed.  The first motion would influence just the starting point. A left motion would shift the starting point towards the decision threshold for a response to the right and vice versa. Here, the drift rate would not be influenced by the first motion. In the middle, altered sensory processing is depicted. The starting point would remain unchanged, the aftereffect would solely be caused by an alteration in the acculumuation process, i.e. a variable drift rate. On the right, the combination of both is presented. The aftereffect would be explained by a shifted starting point as well as a variable drift rate.'}

knitr::include_graphics(path = 'ddm.eps')

```

# Experiment 1

Experiment 1 was originally planed as a modification of the Spatial Cueing paradigm by Posner, Snyder and Davidson [@posner_attention_1980] to examine potential influences of own-body motion stimuli on the perception of subsequent motion stimuli. Similar to Spatial Cueing in the visual system, a cue stimulus was expected to direct covert attention to one side in space.  The spatial cueing effect is usually found for short stimuli with short ISIs [@ruz_review_2002]. Therefore both movements were short (200 ms) and the ISIs ranged from 50 ms to 600 ms.

Participants were first translated in the y-axis (cue) and were then rotated about an earth-vertical axis (target). The first motion was chosen to be stronger compared to the second one to ensure that participants perceive the cue correctly, since this could potentially influence attention. A trial was considered as congruent when both motions had the same spatial direction, e.g. left translations and counter-clockwise rotation. The first motion was chosen to be 3.6 times greater than the intensity of the threshold established in a preliminary study and similar to published data [@valko_vestibular_2012] and the second motion was chosen to be 2.6 times greater than the intensity of the thresholds reported by @grabherr_vestibular_2008.

## Methods

### Participants

Twelve healthy participants participated in the first experiment (four female, mean age 27, range 24-30 years). All participants were right handed according to a German version of the handedness questionnaire by @chapman_measurement_1987. One of the participants received a course credit for participation. None of the participants reported a history of relevant neurological, vestibular or attentional disorders. The study was approved by the ethics committee of the University of Bern and all participants gave written informed consent prior to the experiment in accordance with the Declaration of Helsinki.

### Motion stimuli

A six degree of freedom motion base (6DOF2000E, Moog Inc., East Aurora, NY) and in-house software were used to generate the motion stimuli. Cue (antecedent) stimuli consisted of translations with single cycle sinusoidal acceleration and a frequency $f$ of 5 Hz ($a(t) = A\sin(2 \pi ft) = A\sin(\frac{2 \pi t}{T})$, $T = \frac{1}{f}$) along the y-axis (left/right) and the x-axis (forth/back) as used in previous studies [see e.g., @benson_thresholds_1989; @crane_foreaft_2012; @grabherr_vestibular_2008]. Acceleration amplitude (A) was set to $0.25 \frac{m}{s^2}$, i.e. a peak velocity of $0.016 \frac{m}{s}$ ($v_{max} = \frac{AT}{\pi}$) and a displacement of 0.0016 m ($\Delta p = \frac{AT^2}{2\pi}$). For the target (subsequent) stimuli the same acceleration profile as for the cue stimuli was used but with yaw rotations about an earth-vertical axis (left/right). The acceleration amplitude was set to $24 \frac{deg}{s^2}$ ($v_{max} = 1.53 \frac{deg}{s}, \Delta p = .153 deg$).

### Experimental design

Within each participant Cue Congruence (three levels: congruent, incongruent and neutral) and inter-stimulus interval (ISI, four levels: 50 ms, 100 ms, 200 ms and 600 ms) were manipulated. In congruent trials y-axis translations were followed by yaw rotations to the corresponding side. In incongruent trials y-axis translations preceded yaw rotations to the opposite side. Neutral trials consisted of a translation along the x-axis and a subsequent yaw rotation. The combination of the two factors resulted in 32 motion sets. Every motion set was repeated six times per participant. In-house software based on labVIEW (National Instruments, Austin, Texas) was used to record participants' response and reaction times.

### Experimental procedure

Participants were seated upright in a car seat with a five-point harness which was mounted on the motion base. Their head was centered and fixated with a helmet. The experiments were conducted in the dark and participants were blindfolded to prevent the processing of visual cues. White noise was delivered to in-ear headphones at approximately 60 dB to mask sounds from the motion base. Participants were instructed to indicate the perceived direction of the second motion as fast as possible by pressing the corresponding button in either their left or right hand. In case of uncertainty, participants had to make their best guess. The first motion was described as a time varying warning cue, which would indicate the beginning of the second motion but not predict its direction. Practice trials were administered until participants understood the task. Once a participant was comfortable with the task, motion sets were presented in three blocks. The length of the breaks between blocks were determined by participants. In total, 192 trials were randomly presented (48 congruent, 48 incongruent, 96 neutral). A trial consisted of a cue stimulus (200 ms), an ISI (50-600 ms), a target stimulus (200 ms), time to respond (max. 2500 ms), return to the origin (1100 ms) and an  interval of about 1000 ms.

### Data analysis

For all experiments analyses and figures were computed in R [@R-base] using the ggplot2 [@R-ggplot2], brms [@R-brms] and rstan [@R-rstan] packages. Analyses, models and data are freely accessible on the OpenScienceFramework (OSF) (https://osf.io/46nqw/).

To describe data from experiment 1 in more detail, participants' performances were analyzed with a multilevel Bayesian logistic regression model with factors Cue Congruence (congruent, incongruent, neutral) and ISI (50ms, 100ms, 200ms, 600ms). 
<!---To analyze participants' reaction times a multilevel Bayesian regression model with an ex-Gaussian likelihood function and factors Cue Congruence (congruent, incongruent, neutral) and ISI (50ms, 100ms, 200ms, 600ms) was computed. Only correct trials were included as the underlying mechanism leading to a wrong response might differ. Priors were set to the default implemented in brms.-->

To jointly analyze the participants' response and response times, and to study the bias introduced by the first motion, two different hierarchical Drift Diffusion Models [@vandekerckhove_hierarchical_2011] in a Bayesian framework were calculated in brms [@R-brms]. In a hierarchical model group and participant parameters can be estimated at the same time based on the assumption that participants' parameters are drawn from a group distribution [@vandekerckhove_hierarchical_2011; @wiecki_hddm:_2013]. It allows parameter estimates for groups while taking into account variation between participants. Additionally, the Bayesian implementation allows inference of the complete posterior distributions of the parameters of interest. For the first experiment, neutral conditions were left out to simplify the statistical models.

In the second model, for the drift rate fixed effects of the first direction (left/right), the second direction (clockwise or counter-clockwise) were estimated together with the random effects of participants. Moreover, for the starting point the fixed effects of the first motion and ISI along with the random effect of participants was estimated. For the non-decision time the random effects of participants were estimated.

The better model was selected based on the leave-one-out cross-validation (LOO) [@vehtari_practical_2017].

## Results

```{r figDescriptiveCorrect, fig.cap='The mean of the proportion of correct answers in the direction discrimination task for all participants in all conditions and experiments. The means over all participants are represented by red dots for congruent conditions, green triangles for incongruent conditions and the blue squares for neutral condtions (only experiment 1). The error bars represent the Standard Error. The dotted line marks the chance level.',  echo=FALSE, cache = FALSE, warning=FALSE}
plot_correct
```

```{r figDescriptiveRT, fig.cap='The mean of the proportion of correct answers in the direction discrimination task for all participants in all conditions and experiments. The means over all participants are represented by red dots for congruent conditions, green triangles for incongruent conditions and the blue squares for neutral condtions (only experiment 1). The error bars represent the Standard Error. The dotted line marks the chance level.',  echo=FALSE, cache = FALSE, warning=FALSE}         
rt_times_plot
```

### Accuracy

The logistic regression showed that the motion discrimination performance is below chance when the prior stimulus was in the same direction, e.g. a left translation followed by a counter-clockwise yaw rotation. On the other hand, direction discrimination performance is above chance level when a yaw rotation is preceded by an incongruent or a neutral fore-aft motion (Figure \@ref(fig:figDescriptiveCorrect) - experiment 1). The discrimination performance did not differ meaningfully between the incongruent and neutral motions, as all 95% CIs for the posterior distribution of the different included zero. Overall, the response patterns indicate MAEs, except for ISIs of 600ms.

<!--Moreover, correct response times are faster for incongruent trials (left translation followed by clockwise rotation or right translation followed by counter-clockwise rotation) than for congruent trials (left translation followed by counter-clockwise rotation or right translation followed by counter-clockwise rotation). For incorrect response times the opposite pattern is found. A more detailed analysis for accuracy and reaction times can be found in the supplementary materials. -->

### Drift Diffusion Models

For the first experiment, the winning model included fixed effects of the first direction (left/right), the second direction (clockwise or counter-clockwise), for the drift rate fixed effects of the first direction (left/right), the second direction (clockwise or counter-clockwise) and the random effects of participants (see figure ddm_exp1). To get an estimate for each condition, no intercept and the triple interaction effect was entered.

This model revealed that evidence was accumulated in the direction of the first motion an ISI of 50ms, if the second motion was to the same direction as the first. It appears that evidence is still accumulated for the direction of the first motion when the second motion is to the opposite, yet those 95% credible intervals include zero. Moreover, the second motion did not meaningfully alter evidence accumulation (all CI include zero, with the median around zero).

For the ISI of 100ms evidence is accumulated for the direction of the first motion, except for the condition of the first motion and second motion to the left, where the 95% CI includes zero. Again, the second motion did not meaningfully alter evidence accumulation (all CI include zero, with the median around zero).
For the ISI of 200ms evidence is accumulated for the direction of the first motion, when the second motion is to the other direction. In the congruent conditions, evidence for the first motion direction is accumulated, yet with the 95% CI barely including zero. Here the second motion alters evidence accumulation meaningfully, i.e. it shifted evidence accumulation towards the correct response of the second motion. For the ISI of 600ms, evidence for the second motion is accumulated, if the second motion is to the right. If the second motion is to the left, evidence for the left is accumulated, however in the congruent condition, the 95% CI includes zero.

```{r ddm1Drift, fig.cap='The means for all conditions and their accuracy',  echo=FALSE, cache = FALSE, warning=FALSE}
better_fig_winning_model_exp1_drift 
```

Moreover, the winning model revealed a slightly shifted starting point to the response "right", if the first motion was to the left and vice versa for an ISI of 50ms. However, the 95% credible interval included zero in both starting points (see figure ddm drift exp 1). For an ISI of 100ms, the starting point was shifted towards the threshold of the right response, if the first motion was to the left, with the 95% credibility interval excluding zero. For the same ISI, the starting point was also slightly shifted to the right threshold, if the first motion was to the right, yet with the 95% credibility interval including zero. For, an ISI of 200 and 600ms the bias was slightly shifted towards the left response if the first motion was to the left and to the right response if the first motion was to the right. However, all 95% credible intervals included zero.

```{r ddm1Bias, fig.cap='The means for all conditions and their accuracy',  echo=FALSE, cache = FALSE, warning=FALSE}
better_fig_winning_model_exp1_bias
```

Overall the winning DDM for experiment 1 revealed mainly an altered sensory processing as evidenced by the drift rates and a tendency towards an altered decision making process prior to the evidence accumulation for short ISIs.

## Discussion

In the first experiment it has been shown that there are cross-modal MAEs within the vestibular system in terms of direction discrimination performance if a yaw rotation is preceded by a  translation in the same spatial direction. Thus, a manipulation of the otoliths apparently influence a sensory decision making process in the semicircular canals. However, MAEs are only found for short pauses between the two motion stimuli. More importantly, in a hierarchical DDM the MAEs appear to be explained by a combination of a variable drift and a shifted starting point to the opposite direction of the first motion. It remains to be seen whether for the incongruent conditions whether the correct responses are due to correct sensory evidence accumulation of the second motion or whether the evidence accumulation process is also biased by the first motion. The same response pattern for neutral trials seems to support the former explanation.

# Experiment 2

The results of Experiment 1 led to the question whether the consistent directional misperception of the second motion (motion aftereffect) still exists when a rotation (cue) is preceding a translation (target), more concretely whether the cross-modal MAE found in experiment 1  is bidirectional. The sequence of the two motions was therefore interchanged so that a left/right yaw rotation preceded a left/right translation (while the motion trajectories stayed otherwise the same). Only congruent and incongruent cues were presented, since no rotations could serve as “neutral” cues for lateral translations in the horizontal plane as a pitch or roll rotation in an upright participant would lead inevitably to a tilted/skewed subsequent motion.

## Methods

### Participants

For experiment 2, twelve new participants were recruited. One participant pressed the left button in 92% of all trials and was excluded from the study. The remaining 11 participants (nine female, mean age 25, range 21-38 years). All participants were right handed according to a German version of the handedness questionnaire by @chapman_measurement_1987. None of the participants reported a history of relevant neurological, vestibular or attentional disorders. The study was approved by the ethics committee of the University of Bern and all participants gave written informed consent prior to the experiment in accordance with the Declaration of Helsinki.

### Motion stimuli

Cue stimuli consisted of yaw rotations about an earth-vertical axis (left/right). They consisted of single-cycles of sinusoidal acceleration at a frequency of 5 Hz. Acceleration amplitude was $24 \frac{deg}{s^2}$ ($v_{max} = 1.53 \frac{deg}{s}, \Delta p = .0153 deg$). Target stimuli were translations along the y-axis (left/right). They consisted of the same acceleration profile as the cue stimuli. Frequency was again 5 Hz. Acceleration amplitude was $0.25 \frac{m}{s^2}$ ($v_{max} = .016 \frac{m}{s}, \Delta p = .0016 m$).

### Experimental design

Within each participant Cue Congruence (three levels: congruent, incongruent) and inter-stimulus interval (ISI, four levels: 50 ms, 100 ms, 200 ms and 600 ms) were manipulated. In congruent trials yaw rotation were followed by translations along the y-axis to the corresponding side. In incongruent trials yaw rotations were followed by y-axis translations to the opposite side. The combination of the two factors resulted in 16 motion sets. Every motion set was repeated 12 times per participant. Again, participants' response and reaction times were recorded.

### Experimental procedure

Experimental procedure was similar to the one in Experiment 1. Only the type of motion sets presented was different. In this experiment, 192 motion sets were randomly presented (96 congruent, 96 incongruent).

### Data analysis

To describe data from experiment 2 in more detail, participants' accuracies were analyzed for each experiment with a multilevel Bayesian logistic regression model factors Cue Congruence (congruent, incongruent) and ISI (50ms, 100ms, 200ms, 600ms).

<!--For the reaction times a Bayesian multilevel regression with a ex-Gaussian likelihood function and factors Cue Congruence (congruent, incongruent) and ISI (50ms, 100ms, 200ms, 600ms) was computed. Only correct trials were included as the underlying mechanism leading to a wrong response might differ. Priors were set to the default implemented in brms.-->

To jointly analyze the participants' response and response times, and to study the bias introduced by the first motion, two different hierarchical Drift Diffusion Models [@vandekerckhove_hierarchical_2011] were calculated in brms [@R-brms]. Neutral conditions were eliminated from the data set to simplify the statistical models.

In the first model, for the drift rate fixed effects of the first direction (counter-clockwise/clockwise), the second direction were estimated together with the random effects of participants. For the non-decision time and starting point the random effects of participants were estimated.

In the second model, for the drift rate fixed effects of the first direction (counter-clockwise/clockwise), the second direction (left/right) were estimated together with the random effects of participants. Moreover, for the starting point the fixed effects of the first motion and ISI along with the random effect of participants was estimated. For the non-decision time the random effects of participants were estimated.

The better model was selected based on the leave-one-out cross-validation (LOO) [@vehtari_practical_2017].

## Results

### Accuracy

The logistic regression model (In Fig accuracy plot Experiment 2) shows that the motion discrimination performance is around (for ISIs of 50 and 100 ms) or even above chance level (for ISIs 200 and 600ms) when the prior stimulus was congruent with the motion, e.g. a counter-clockwise yaw rotation followed by a left translation. Direction discrimination performance for the second motion is higher when a yaw rotation is followed by an incongruent  translation. In this Experiment there are no clear MAEs.

### Drift Diffusion Models
For the second experiment, the winning model included fixed effects of the first direction (left/right), the second direction (clockwise or counter-clockwise), for the drift rate fixed effects of the first direction (left/right), the second direction (clockwise or counter-clockwise) and the random effects of participants (see figure ddm_exp1).

This model revealed that evidence was accumulated in the direction of the second motion for an ISI of 50, 100 and 200ms, if the second motion was to the opposite direction. If it was to the same direction, evidence for the second motion was accumulated, however with the 95% CI including zero. same direction as the first. It appears that evidence is still accumulated for the direction of the first motion when the second motion is to the opposite, yet those 95% credible intervals include zero. The second motion meaningfully shifted the evidence accumulation process to direction of the correct response. 
For the ISI of 600ms, evidence for the second motion was always accumulated accumulated. The second motion did not meaningfully alter evidence accumulation (all CI include zero, with the median around zero).
Moreover, the winning model revealed no shift of starting point altered as all 95% CI included zero.

```{r ddm2Drift, fig.cap='The means for all conditions and their accuracy',  echo=FALSE, cache = FALSE, warning=FALSE}
better_fig_winning_model_exp2_drift 
```

```{r ddm2Bias, fig.cap='The means for all conditions and their accuracy',  echo=FALSE, cache = FALSE, warning=FALSE}
better_fig_winning_model_exp2_bias
```

## Discussion

In experiment 2, no MAEs were found although participants' performance in congruent trials was still worse than in incongruent trials and was at chance level. In experiment 2 motion stimuli were the same as in experiment 1, simply in a different order. The cue stimuli in Experiment 1 were stronger in terms of intensity compared to the target stimuli. The switch of the order was switched without adapting of the intensities could explain the absence of MAEs.

# Experiment 3

In experiment 3 it was hypothesized that the MAE would reappear if the intensities of the stimuli in terms of distance to thresholds were in the same ratio as in experiment 1. For this third experiment the rotation about an earth-vertical axis was chosen at an intensity 3.6 times above the threshold found by @grabherr_vestibular_2008 and the horizontal translation was chosen at an intensity 2.6 times above the threshold found by Lim and colleagues (unpublished data).

## Methods

### Participants
Twelve new participants were recruited for the third experiment. Two participants had to abort the experiment. The remaining ten participants (four female, mean age 29, range 23-56 years). All participants were right handed according to a German version of the handedness questionnaire by @chapman_measurement_1987. None of the participants reported a history of relevant neurological, vestibular or attentional disorders. The study was approved by the ethics committee of the University of Bern and all participants gave written informed consent prior to the experiment in accordance with the Declaration of Helsinki.

### Motion stimuli

Cue stimuli were rotations about an earth-vertical axis (right/left) with single cycle sinusoidal acceleration and a frequency of 5 Hz ($a(t) = A \sin (2 \pi ft) = A \sin (\frac{2\pi t}T)$). Acceleration amplitude was set to $33 \frac{deg}{s^2}$ ($v_{max} = 2.10 \frac{deg}{s}, \Delta p = 0.021 deg$). Target stimuli consisted of translations in the y-axis (right/left) and the same acceleration profile as the cue stimuli. Acceleration amplitude was set to $0.18 \frac{m}{s^2}$ ($v_{max} = .011 \frac{m}{s}, \Delta p = .0011 m$).

### Experimental design and procedure

Design and experimental procedure were identical to Experiment 2.

### Data analysis

To describe data from experiment 2 in more detail, participants' accuracies were analyzed for each experiment with a multilevel Bayesian logistic regression model factors Cue Congruence (congruent, incongruent) and ISI (50ms, 100ms, 200ms, 600ms).

<!--For the reaction times a Bayesian multilevel regression with a ex-Gaussian likelihood function and factors Cue Congruence (congruent, incongruent) and ISI (50ms, 100ms, 200ms, 600ms) was computed. Only correct trials were included as the underlying mechanism leading to a wrong response might differ. Priors were set to the default implemented in brms.-->

To jointly analyze the participants' response and response times, and to study the bias introduced by the first motion, two different hierarchical Drift Diffusion Models [@vandekerckhove_hierarchical_2011] were calculated in brms [@R-brms]. Neutral conditions were eliminated from the data set to simplify the statistical models.

In the first model, for the drift rate fixed effects of the first direction (counter-clockwise/clockwise), the second direction were estimated together with the random effects of participants. For the non-decision time and starting point the random effects of participants were estimated.

In the second model, for the drift rate fixed effects of the first direction (counter-clockwise/clockwise), the second direction (left/right) were estimated together with the random effects of participants. Moreover, for the starting point the fixed effects of the first motion and ISI along with the random effect of participants was estimated. For the non-decision time the random effects of participants were estimated.

The better model was selected based on the leave-one-out cross-validation (LOO) [@vehtari_practical_2017].

## Results

### Accuracy

The logistic regression model (In Fig Acc Experiment 3) revealed that the motion discrimination performance is below chance when the prior stimulus was in the same direction, e.g. a left translation followed by a counter-clockwise yaw rotation for ISIs of 50, 100 and 200 ms. For an ISI of 600ms the discrimination performance is around chance. Direction discrimination performance is above chance level when a yaw rotation is preceded by an incongruent or fore-aft motion. As in experiment 1, the response patterns indicate MAEs, except for ISIs of 600ms.

<!--RESPONSE TIMES STILL OPEN: Again, correct response times are faster for incongruent trials (left translation followed by clockwise rotation or right translation followed by counter-clockwise rotation) than for congruent trials (left translation followed by counter-clockwise rotation or right translation followed by counter-clockwise rotation), while the opposite pattern is found for incorrect response times. A more detailed analysis for accuracy and reaction times can be found in the supplementary materials. -->

### Drift Diffusion Model

Based on the LOO criterion, the selected model included fixed effects of the first motion, the second motion, and ISI on the drift rate. The starting point was allowed to vary as a function of the of the first motion and ISI (fixed effects), as well as the random effects of participants (see figure ddm_exp1). 
This model revealed that evidence was accumulated in the direction of the first motion for all ISIs except for congruent motions at an easy of 600 ms (see figure Drift Rate, Experiment 3). The second motion did not meaningfully alter evidence accumulation for the ISIs of 50-200, as the 95% credibility intervals of the posterior distributions of the differences within the first motions include zero (see table DIFF). For the congruent conditions and an ISI of 600 ms, the second motion appears to shift the drift rate to the correct response of the second motion, although the medians (???) still point towards evidence accumulation for the direction of the first motion.

```{r ddm3Drift, fig.cap='The means for all conditions and their accuracy',  echo=FALSE, cache = FALSE, warning=FALSE}
better_fig_winning_model_exp3_drift 
```

Moreover, the winning model revealed a starting point shifted to the response "right", if the first motion was to the left and the ISI 50ms. For the shortest ISI an a motion to the right, the bias was slightly shifted towards the threshold of the left response, but the 95% credibility interval included 0. The opposite is seen for an ISI of 100ms and the first motion to the left. For an ISI of 100 ms and a motion to the right as well as an ISI of 200 ms and a motion to the left, there was no shift of the starting point. However, for an ISI of 200ms and the first motion to the right, the starting point was shifted towards the threshold of the "right" response. The same pattern is found for an ISI of 600ms. There, the bias was shifted towards the response threshold "left", if the first motion was in the left direction and shifted to the threshold "right", if the first motion was in the right direction. 
Overall the winning DDM for experiment 3 revealed a combination of altered sensory processing as evidenced by the drift rates and an altered decision making process prior to the evidence accumulation as shown by the shifted starting points for short ISIs.

```{r ddm3Bias, fig.cap='The means for all conditions and their accuracy',  echo=FALSE, cache = FALSE, warning=FALSE}
better_fig_winning_model_exp3_bias
```

## Discussion

Performance data clearly indicate MAEs for short ISI even stronger than in experiment 1. Thus, in experiment 3 it has been shown that the cross-modal MAE within the vestibular system are also present when a yaw translation is followed by a  translation in the same spatial direction. Thus, cross-modal vestibular MAEs appear to be bidirectional. The analysis of the hierarchical DDM has shown that the MAE's underlying decision making process could be explained by a combination of a variable drift and a shifted starting point to the opposite direction of the first motion. Compared to experiment 2, the intensity ratio of the first motion (yaw rotation) to the second motion (translation) was adapted to be similar to the  intensity ratio of experiment 1 (first translation than rotation).

# General Discussion
In this three spatial cueing experiments were used to reanalyze the data in the light of recently discovered vestibular MAEs [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012]. In two out of three experiments, we replicated vestibular MAEs [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012] and showed that they cross modal MAEs exist even between two different vestibular receptors. In the the first experiment it has been shown that strong translations to the left or to the right drastically influence the decision making process of yaw rotation to the same spatial direction. These MAEs were also observed in experiment 3 when a yaw rotation preceded a translation to the same spatial direction. These results are striking as participants systematically misjudged strong motion stimuli well above the detection threshold as a function of the first motion. Classically, visual MAEs are observed in stationary patterns after a motion exposure. Generally, MAEs are limited to object motion. Here, we repliacted that they also apply to self-motion perception.

In experiment 2 although the direction discrimination performance was still worse when the second motion was preceded by a motion in the same spatial direction, there were no prototypical MAEs, i.e. the perception of the motion to the opposite direction. Experiment 2 and 3 solely differed in the ratio of the intensity of the first and second motion. In experiment 2 the second motion was 3.6 and the first motion 2.6 times above the intensity thresholds [@grabherr_vestibular_2008; @valko_vestibular_2012], while in experiment 3 the opposite intensity ratios were used. Although we cannot directly compare the intensity of the yaw rotation and  translation, it seems that the first motion's intensity has to be stronger relative to the intensity threshold than the second motion. It is unclear whether this finding is generalizable as in experiment 1 the second motion's intensity was always relatively weaker. Additionally, uni-modal vestibular MAEs [@coniglio_human_2014; @crane_foreaft_2012; @crane_roll_2012] and cross-modal visuo-vestibular MAEs [@cuturi_optic_2014] are usually shown with an adaptive staircase paradigm. While the intensity and duration of the first motion is always fixed, the second motion's intensity takes into account the participants' previous response separately for each direction of the first motion. It is therefore difficult to draw conclusions about the needed intensity ratios between the first and second motion in order to produce MAEs.

Another crucial factor for MAEs are the temporal intervals between two motions. In the present study we were limited to the ISIs selected for the Spatial Cueing Paradigm ranging from 50 to 600ms. Compared to the literature on MAEs those ISIs are rather short. For example @crane_foreaft_2012 used ISIs from 500 to 3000ms and only found consistent MAEs starting at 1000ms. Similarly, @coniglio_human_2014 observed weaker MAEs with increasing ISIs. In contrast to those studies, in the present study MAEs were only present with ISIs of 50, 100 and 200ms. In the visual modality at such short ISIs (<500ms) priming effects have been observed. A priming effect describes a facilitated perception of the motion in the same direction as the first one [@kanai_perceptual_2005].

Also, the presentation duration of the cue stimulus in the present study (200ms) was much shorter than the adapter stimulus in @crane_foreaft_2012 (1500ms). Again in the visual domain, MAEs have been observed after an adaption of only 320ms [@kanai_perceptual_2005]. On the other hand, simulations of 80ms lead to a priming effect. However, the direct comparison of temporal aspects of perception between the vestibular and visual systems may not be advisable as vestibular signals are perceived slowly (COWAN). On the other side of the temporal spectrum, it has been known for a long time that a prolonged yaw rotation leads to a continuation of motion perception in the same direction once the motion has stopped [e.g. @bertolini_velocity_2011]. This illusion is driven the neural pathways that control eye movements and is known as Velocity Storage. In contrast, a prolonged motion stimulation of visual motion most certainly results in MAEs [@anstis_motion_1998]. It remains to be seen whether so called rapid MAEs, using short stimuli, and the classical MAE, where a prolonged stimulation is used, are both based on the same neural adaptation mechanism [@anstis_motion_1998].

The design of the experiments was not intended to find temporal aspects of MAEs as it was customized to Spatial Cue Paradigm. Therefore, it is difficult to go beyond speculative remarks in this aspect. The present data combined with previous studies could suggest the appearance of a MAE as a function of the interactions of stimulus duration, stimulus intensity and ISI. This idea might be supported by the observation that a long stimulus intensity is less affected by longer ISIs than short stimuli [@kanai_perceptual_2005]. Future studies investigating the vestibular MAE should systematically vary those parameters in order to detect the limits of MAEs similarly to @kanai_perceptual_2005 in the visual modality. Moreover, the used paradigm differ substantially from what was normally used to observe (vestibular) MAEs. This makes it difficult to directly compare the parameter specifics of the current studies with those of previous ones. Nevertheless, the decision to use constant and not adaptive stimuli nicely enrich literature on MAE.

The cross-modal MAE within the vestibular system observed in this study is highly interesting for several reasons. It suggests that counter-clockwise rotation and translations to the left, and clockwise rotations and translations to the right share their representation at some level. While this makes intuitive sense, it is not obvious on a more abstract level, at least we do not know any literature (IS THERE LITERATURE?). If neural adaptation is the explanation for the present MAE, it could mean that no distinction is made between translation and rotation at some (neural) level. From the current results it is difficult to conclude at which processing state of perception vestibular MAEs take place as the signals of the semicircular canals and the otoliths are integrated early on in the vestibular nuclei [e.g., @carriot_integration_2015; @dickman_vestibular_2002]. The addition of response times in the present study allowed us to fit a DDM, a cognitive processing model, in order to understand the perceptual decision making process better. In fact, the drift rate changing as a function of the first and second motion shows that the first motion alters sensory evidence accumulation and thus dynamic sensory processing. Based on our data we cannot pinpoint the location but only speculate. @kanai_perceptual_2005 argue that short stimuli should evoke a MAE at very early processing stages, in their case the primary visual areas. However, next to the drift rate varying as a function of the first motion, the DDM also reveled a shift of the starting point in spatially congruent conditions towards the opposite direction. This indicates a change in the perceptual decision making process prior to the actual sensory evidence accumulation and can be interpreted as a higher level alteration, possibly independent of the sensory accumulation process [@ellis_vestibular_2017; @wiech_influence_2014]. Those findings fuel the idea to manipulate the MAEs with more cognitive factors such as systematically varied prior information [@ellis_vestibular_2017]. Such studies could possibly show whether it possible to alter MAEs, considered a lower level phenomenon, with higher level strategies.

Another important aspect rarely touched in vestibular studies that make use of passive vestibular stimulation is the influence of proprioception. Participants are usually seated and fixed on a motion platform. Accelerations and decelerations not only stimulate the vestibular system but also generate relevant proprioceptive and tactile information that are hardly distinguishable from the pure vestibular input. In the present study rather strong stimuli were used and therefore, it cannot be excluded that proprioceptive input contributed to the MAE as well given the fact that MAEs have been reported for proprioceptive [@seizova-cajic_proprioceptive_2007] and visuo-tactile manipulations [@konkle_motion_2009]. Thus, the reported strong MAEs might actually be cross-modal including the otoliths, semicircular canals and proprioception.

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}